---
title: "Reproducible Research First Assignment"
author: "Khaled Alzafari"
date: "September 16, 2015"
output: word_document
---

## load the Data:
```{r}
echo = TRUE
mydata= read.csv(file="activity.csv", header=TRUE, sep=",")
```


## Histogram for the total number of steps per day:
```{r}
echo = TRUE
totalSteps <- aggregate(steps ~ date, data = mydata, sum, na.rm = TRUE)
library(ggplot2)
ggplot(data=totalSteps, aes(totalSteps$steps)) + geom_histogram( col="black", fill="red", alpha = 0.3, binwidth=1500) + labs(title="Histogram for Total Steps Per Day") + labs(x="Total Steps", y="Count") + geom_density(alpha=0.9)

```

## the mean and median total number of steps taken per day:

```{r}
echo = TRUE
Mean_steps = mean(totalSteps$steps,  na.rm = TRUE)
Median_steps = median(totalSteps$steps,  na.rm = TRUE)
Mean_steps
Median_steps
```

## The average daily activity pattern: 

```{r}
echo = TRUE
meanSteps <- aggregate(steps ~ interval, data = mydata, mean, na.rm = TRUE)

dp= ggplot(meanSteps, aes(meanSteps$interval, meanSteps$steps))+ geom_line(col="red", fill="red", alpha = 0.5) 
dp + labs(title="Time seires for Average number of steps Taken at 5 minute Intervals") + labs(x="T5-min interval", y="average number of steps taken")

```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
echo = TRUE
maxinterval = meanSteps[which.max(meanSteps$steps), ]
maxinterval
### It is the interval number 835  that has the value of 206.2
```

##Imputing missing values
### Number of Missing Value

```{r}
echo = TRUE
missing_rows <- sum(!complete.cases(mydata))
missing_rows
```

## Devise a strategy for filling in all of the missing values in the dataset the new stratgy depend on replace NA with the mean of the "steps" variable
```{r}
echo = TRUE
newdataset= mydata
newdataset$steps[which(is.na(newdataset$steps))] = mean(newdataset$steps, na.rm = T)
```

## histogram for the new Data Set:
```{r}
echo = TRUE
newtotalSteps <- aggregate(steps ~ date, data = newdataset, sum, na.rm = TRUE)
library(ggplot2)
ggplot(data=newtotalSteps, aes(newtotalSteps$steps)) + geom_histogram( col="black", fill="blue", alpha = 0.3, binwidth=1500) + labs(title="Histogram for Total Steps Per Day") + labs(x="Total Steps", y="Count") + geom_density(alpha=0.9)
```

## The mean and median total number of steps taken per day for the new data set:
```{r}
echo = TRUE
newMean_steps = mean(newtotalSteps$steps)
newMedian_steps = median(newtotalSteps$steps)
newMean_steps
newMedian_steps

```

## What is the impact of imputing missing data on the estimates of the total daily number of steps?

### The mean for both data (before and after removing missing Data NA) while our strategy was based on replacing NA with the mean of the same variable.whereas, the there is a small difference for the median before removeing NA and after. 


## Are there differences in activity patterns between weekdays and weekends?

```{r}
echo = TRUE
newdataset$days= weekdays(as.Date(newdataset$date))
library(plyr)
newdataset$weeks <- revalue(newdataset$days, c("Saturday"="weekend", "Sunday"="weekend", "Monday"="weekday", "Tuesday"="weekday", "Wednesday"="weekday","Thursday"="weekday", "Friday"="weekday" ))
newmeanSteps <- aggregate(steps ~ interval + weeks, data = newdataset, mean, na.rm = TRUE)

library(ggplot2)
X1 = ggplot (newmeanSteps, aes(newmeanSteps$interval, newmeanSteps$steps, colour = newmeanSteps$weeks) ) + geom_line() 
X1 + labs(title="Average number of steps Taken at 5 minute Intervals") + labs(x="T5-min interval", y="average number of steps taken") + labs(colour="Type of week Day")
```

### The difference is clear between the normal working days and weekend days...